---
title: "Coastal Copula Example"
format: html
editor: visual
author: Avital Breverman
---

This code builds two conditional extreme-flow dependence models (Scenario A/B), selects copulas by AIC, generates joint extreme simulations, derives joint return-period contours, merges them into a conservative envelope, and then ranks where along that envelope joint outcomes are most probable using KDE—yielding both curves and a representative “most likely” design point.

Here is a summary of the steps:

-   Import data from CSV

-   Extract a partial duration series at each location

-   Compute rank correlation coefficient (Kendall's tau) between the two locations

-   Define copula scenarios

-   Fit copula models and select copula for each scenario based on Akaike Information Criterion (AIC)

-   Generate samples from each copula and transform to original variable scale

-   Derive isoline for specified probabilities (return periods)

-   Combine isolines into envelope curve

-   Compute relative density of pairs on each isoline using kernel density estimation

```{r}
rm(list = ls())

# install.packages("remotes")
# remotes::install_github("rjaneUCF/MultiHazard")

library(tidyverse)
library(POT)
library(lmom)
library(lmomco)
library(VineCopula)
library(MultiHazard)
library(purrr)
library(geomtextpath)
```

# Functions

Define custom functions. There are additional functions in later code blocks. I'm sorry for my R sins.

```{r}
# Madsen annualization of Generalized Pareto distribution to Generalized Extreme Value distribution 
# gpd_par: Generalized Pareto distribution parameters
# lambda: rate of events
# returns Generalized Extreme Value (GEV) distribution parameters
# Reference: Madsen, H., Rasmussen, P., & Rosbjerg, D. (1997). Comparison of annual maximum series and partial duration series methods for modeling extreme hydrologic events 1. At-site modeling. Water Resources Research, # 33(4), 747-757.
madsen_annualization <- function(gpd_par, lambda) {
  gev_xi <- gpd_par["xi"] + (gpd_par["alpha"] / gpd_par["k"]) * (1 - lambda ^ -gpd_par["k"]) # scale
  gev_alpha <- gpd_par["alpha"] * lambda ^ -gpd_par["k"] # scale
  gev_k <- gpd_par["k"]  
  
  print(gev_xi)
  print(gev_alpha)
  print(gev_k)
  
  c(xi = as.numeric(gev_xi),
  alpha = as.numeric(gev_alpha),
  k = as.numeric(gev_k))
}

# Langbein adjustment defines the relationship between empirical flood expectancies in the partial duration series and the probability of the correspoding flood as an annual series
# p_pds: partial duration non-exceedance probability
# lambda: rate of events
# returns the annual non-exceedance probability
# Reference: Langbein, W. B. (1949). Annual Floods and the Partial-duration Flood Series. American Geophysical Union, 30(6), 879-881.
langbein_adjustment <- function(lambda, p_pds) {
  exp(-lambda * (1 - p_pds))
}
```

# Import Data from CSV

```{r}
surge <- read.csv("RG_Sg_2022.csv", header = TRUE, sep = ",", na.strings = c("-9999")) %>% mutate(Value = Surge,                                                                     Date = as.POSIXct(dstring, format = "%m/%d/%Y %H:%M", tz = "UTC")) %>% select(Date, Value)

precip <- read.csv("Rg_Pr_2022_v2.csv", header = TRUE, sep = ",") %>% mutate(Value = PR_data_mm, Date = as.POSIXct(Pr_Dates, format = "%m/%d/%Y %H:%M", tz = "UTC")) %>% select(Date, Value)

# User-specified variable names and plot labels
surge_name <- "Surge"
precip_name <- "Precipitation"

surge_label <- "Surge (m)"
precip_label <- "Precipitation (mm)"

scenario_A_label <- "Scenario A: Extremal Surge, Coincident Precipitation"
scenario_B_label <- "Scenario B: Extremal Precipitation, Coincident Surge"
```

# Plot Data

```{r}
surge_time_series_plot <- ggplot(surge, aes(x = Date, y = Value)) +
  geom_line() +
  theme_bw() +
  xlab("") +
  ylab(surge_label) +
  ggtitle(surge_name)
surge_time_series_plot

precip_time_series_plot <- ggplot(precip, aes(x = Date, y = Value)) +
  geom_line() +
  theme_bw() +
  xlab("") +
  ylab(precip_label) +
  ggtitle(precip_name)
precip_time_series_plot
```

# Extract Surge PDS

```{r}
surge_clean <- surge |> filter(!is.na(Date) & !is.na(Value))
surge_clean$Date  <- with_tz(surge_clean$Date,  "UTC")
# X_clean$Date  <- round_date(X_clean$Date,  unit = "hour")
surge_clean <- surge_clean %>%
  group_by(Date) %>%
  slice_max(Value, n = 1, with_ties = FALSE) %>%
  ungroup()

# Create data frame with obs and time columns
surge_pot <- data.frame(
  obs = surge_clean$Value,
  time = as.numeric(surge_clean$Date)  # Convert to numeric
)

# Remove any rows with NA, NaN, or Inf
surge_pot <- surge_pot[complete.cases(surge_pot), ]
surge_pot <- surge_pot[is.finite(surge_pot$obs) & is.finite(surge_pot$time), ]

time_sep <- 3600 * 24 * 2 # 2 days in seconds
thresh_surge <- 0.1581 # m

# Extract partial duration series
pds_surge <- POT::clust(data = surge_pot,
                              u = thresh_surge, # threshold
                              tim.cond = time_sep, # time separation
                              clust.max = TRUE) # extract cluster maxima

summary(pds_surge)
nrow(pds_surge)

# Build dataframe
df_pds_surge <- data.frame(pds_surge) %>%
  rename(Date = time,
         Value = obs) %>%
  mutate(Date = as.POSIXct(Date, format = "%m/%d/%Y %H:%M", tz = "UTC"))  %>%
  select(Date, Value)

df_pds_surge$Date <- with_tz(df_pds_surge$Date, "UTC")
df_pds_surge$Date <- round_date(df_pds_surge$Date, unit = "hour")

# Compute rate of events
years_surge <-
  as.numeric((
    max(surge_clean$Date) - min(surge_clean$Date) + 1
  ) / 365.25)
rate_surge <- nrow(df_pds_surge) / years_surge

# Plot original time series and PDS
surge_time_series_plot +
  geom_hline(yintercept = thresh_surge, linetype = "dashed") +
  geom_point(data = df_pds_surge, aes(x = Date, y = Value), colour = "#A60F2D", shape = 19, size = 2) 
```

# Extract Precipitation PDS

```{r}
precip_clean <- precip |> filter(!is.na(Date) & !is.na(Value))
precip_clean$Date  <- with_tz(precip_clean$Date,  "UTC")
# Y_clean$Date  <- round_date(Y_clean$Date,  unit = "hour")
precip_clean <- precip_clean %>%
  group_by(Date) %>%
  slice_max(Value, n = 1, with_ties = FALSE) %>%
  ungroup()

# Create data frame with obs and time columns
pot_precip <- data.frame(
  obs = precip_clean$Value,
  time = as.numeric(precip_clean$Date)  # Convert to numeric
)

# Remove any rows with NA, NaN, or Inf
pot_precip <- pot_precip[complete.cases(pot_precip), ]
pot_precip <- pot_precip[is.finite(pot_precip$obs) & is.finite(pot_precip$time), ]

# Verify it's clean
anyNA(pot_precip)
any(!is.finite(as.matrix(pot_precip)))

time_sep <- 3600 * 24 * 2 # 3 days in seconds
thresh_precip <- 10 # mm

# Extract partial duration series
pds_precip <- POT::clust(data = pot_precip,
                              u = thresh_precip, # threshold
                              tim.cond = time_sep, # time separation
                              clust.max = TRUE) # extract cluster maxima

summary(pds_precip)
nrow(pds_precip)

df_pds_precip <- data.frame(pds_precip) %>%
  rename(Date = time,
         Value = obs) %>%
  mutate(Date = as.POSIXct(Date, format = "%m/%d/%Y %H:%M", tz = "UTC"))  %>%
  select(Date, Value)

df_pds_precip$Date <- with_tz(df_pds_precip$Date, "UTC")
df_pds_precip$Date <- round_date(df_pds_precip$Date, unit = "hour")

# Compute rate of events
years_precip <-
  as.numeric((
    max(precip_clean$Date) - min(precip_clean$Date) + 1
  ) / 365.25)
rate_precip <- nrow(df_pds_precip) / years_precip

# Plot original time series and PDS
precip_time_series_plot +
  geom_hline(yintercept = thresh_precip, linetype = "dashed") +
  geom_point(data = df_pds_precip, aes(x = Date, y = Value), colour = "#A60F2D", shape = 19, size = 2) 
```

# Rank Correlation: Kendall's tau

Compute rank correlation between variables X and Y using Kendall's rank correlation coefficient (Kendall's tau) over a range of lags.

I computed Kendall's tau under 2 scenarios:

-   Rank correlation between the PDS of variable X (surge) and lagged variable Y (precipitation)

-   Rank correlation between the PDS of variable Y (precipitation) and lagged variable X (surge)

```{r}

lag <- 12 * 24 # -12 to +12 day lags

surge_con <- left_join(df_pds_surge, precip_clean, by = "Date") %>%
  select(Date, Value.x, Value.y) %>%
  rename(surge = Value.x, precip = Value.y)

head(surge_con)
nrow(surge_con)

# Calculate Kendall's tau at different lags
result <- MultiHazard::Kendall_Lag(Data = surge_con,
                      Lags = seq(-lag, lag, 1),  # lag is applied to second named variable
                      PLOT = TRUE, # display plot
                      GAP = 0.1) # length of y-axis above and below max and min                                       

# Extract results
tau_values <- result$Value
p_values <- result$Test

# View tau coefficients
print(tau_values)
print(p_values)


#df_pds_Y$Date <- as.POSIXct(df_pds_Y$Date, format = "%Y-%m-%d %H:%M:%S", tz = "UTC")
#X_clean$Date  <- as.POSIXct(X_clean$Date,  format = "%Y-%m-%d %H:%M:%S", tz = "UTC")

precip_con <- left_join(df_pds_precip, surge_clean, by = "Date") %>%
  na.omit() %>%
  select(Date, Value.x, Value.y) %>%
  rename(precip = Value.x,
         surge = Value.y) 

head(precip_con)
nrow(precip_con)

# Calculate Kendall's tau at different lags
result <- MultiHazard::Kendall_Lag(Data = precip_con,
                      Lags = seq(-lag, lag, 1),  # lag is applied to second named variable
                      PLOT = TRUE, # display plot
                      GAP = 0.1) # length of y-axis above and below max and min                                       

# Extract results
tau_values <- result$Value
p_values <- result$Test

# View tau coefficients
print(tau_values)
print(p_values)
```

The maximum tau in both scenarios occurs with a lag of 0, meaning that the coincident events should be pulled from the same date as the extremal (PDS) events.

# Scenario A (Extremal Surge) Marginal Distributions

Scenario A is defined by extremal surge events (PDS) and the coincident precipitation events. The conditioning variable is surge which is conditioned on precipitation.

```{r}

ggplot(surge_con, aes(x = surge, y = precip)) +
    geom_point() +
    xlab(surge_label) + 
    ylab(precip_label) +
    ggtitle("Scenario A: Extremal Surge, Coincident Precipitation") +
    theme_bw()

# Fit marginal distributions
surge_gpa <- lmom::pelgpa(lmom::samlmu(surge_con$surge)) 
surge_gev <- madsen_annualization(surge_gpa, rate_surge)
precip_glo <- lmom::pelglo(lmom::samlmu(surge_con$precip))

evplot(surge_con$surge, main = surge_label); 
#evdistq(quagev, para = surge_gev)

surge_con_gev <- surge_con %>%
  mutate(surge_cdf = lmom::cdfgev(surge, surge_gev),
         precip_cdf = lmom::cdfglo(precip, precip_glo))

surge_con_gpa <- surge_con %>%
  mutate(surge_cdf = lmom::cdfgpa(surge, surge_gpa),
         precip_cdf = lmom::cdfglo(precip, precip_glo))
```

# Scenario B (Extremal Precipitation) Marginal Distributions

Scenario A is defined by extremal precipitation events (PDS) and the coincident surge events. The conditioning variable is precipitation which is conditioned on surge.

```{r}

ggplot(precip_con, aes(x = precip, y = surge)) +
    geom_point() +
    xlab(precip_label) + 
    ylab(surge_label) +
    ggtitle("Scenario B: Extremal Precipitation, Coincident Surge") +
    theme_bw()

# Fit marginal distributions
precip_gpa <- lmom::pelgpa(lmom::samlmu(precip_con$precip)) 
precip_gev <- madsen_annualization(precip_gpa, rate_precip)
surge_glo <- lmom::pelglo(lmom::samlmu(precip_con$surge))

evplot(precip_con$precip, main = precip_label); 
#evdistq(quagev, para = precip_gev)

precip_con_gev <- precip_con %>%
  mutate(precip_cdf = lmom::cdfgev(precip, precip_gev),
         surge_cdf = lmom::cdfglo(surge, surge))

precip_con_gpa <- precip_con %>%
  mutate(precip_cdf = lmom::cdfgpa(precip, precip_gpa),
         surge_cdf = lmom::cdfglo(surge, surge))
```

# Plot Coincident Events for Each Scenario

```{r}

surge_con_longer <- surge_con |>
pivot_longer(cols = -Date, names_to = "series", values_to = "Variable")

ggplot(X_con_longer, aes(x = Date, y = Gage, color = series)) +
  geom_point(size = 2) +
    geom_hline(yintercept = thresh_X, color = "#A60F2D", linetype = "solid") +
    labs(color = "Variable") + 
  scale_color_manual(                            
    values = c(
      X = "#A60F2D",
      Y = "black")) +
  ylab("") +
  ggtitle("Scenario A: Extremal Surge, Coincident Precipitation") +
  theme_bw()

y_con_longer <- y_con |>
  pivot_longer(cols = -Date, names_to = "series", values_to = "Gage")

ggplot(y_con_longer, aes(x = Date, y = Gage, color = series)) +
  geom_point(size = 2) +
    geom_hline(yintercept = thresh_Y, color = "#A60F2D", linetype = "solid") +
    labs(color = "Gage") +
  scale_color_manual(                            
    values = c(
      Y = "#A60F2D",
      X = "black")) +
  ylab("") +
  ggtitle("Scenario B: Extremal Precipitation, Coincident Surge") +
  theme_bw()
```

# Write Datasets to csv

This code block is for Neptune.

```{r}
write.csv(X_con, file = "X_con.csv")
write.csv(Y_con, file = "Y_con.csv")
```

# Scenario A Copula Fits and Comparisons

```{r}

u <- pull(X_con_gpa, x_cdf)
v <- pull(X_con_gpa, y_cdf)

gumbel_cop <- BiCopEst(u, v, family = 4)
print(gumbel_cop$AIC)
plot(gumbel_cop, type = "contour", margins = "unif", main = "Gumbel", col = "blue")
points(x=u, y=v, pch = 16)

t_cop <- BiCopEst(u, v, family = 2)
print(t_cop$AIC)
plot(t_cop, type = "contour", margins = "unif", main = "Student's t", col = "blue")
points(x=u, y=v, pch = 16)

gauss_cop <- BiCopEst(u, v, family = 1)
print(gauss_cop$AIC)
plot(gauss_cop, type = "contour", margins = "unif", main = "Gaussian", col = "blue")
points(x=u, y=v, pch = 16)

clayton_cop <- BiCopEst(u, v, family = 3)
print(clayton_cop$AIC)
plot(clayton_cop, type = "contour", margins = "unif", main = "Clayton", col = "blue")
points(x=u, y=v, pch = 16)

frank_cop <- BiCopEst(u, v, family = 5)
print(frank_cop$AIC)
plot(frank_cop, type = "contour", margins = "unif", main = "Frank", col = "blue")
points(x=u, y=v, pch = 16)

joe_cop <- BiCopEst(u, v, family = 6)
print(joe_cop$AIC)
plot(joe_cop, type = "contour", margins = "unif", main = "Joe", col = "blue")
points(x=u, y=v, pch = 16)

# Create summary table
copula_tbl <- dplyr::bind_rows(
  tibble(Copula = "Gumbel",       family = gumbel_cop$family,  par = gumbel_cop$par,  par2 = gumbel_cop$par2,  AIC = gumbel_cop$AIC),
  tibble(Copula = "Student's t",  family = t_cop$family,       par = t_cop$par,       par2 = t_cop$par2,       AIC = t_cop$AIC),
  tibble(Copula = "Gaussian",     family = gauss_cop$family,   par = gauss_cop$par,   par2 = gauss_cop$par2,   AIC = gauss_cop$AIC),
  tibble(Copula = "Clayton",      family = clayton_cop$family, par = clayton_cop$par, par2 = clayton_cop$par2, AIC = clayton_cop$AIC),
  tibble(Copula = "Frank",        family = frank_cop$family,   par = frank_cop$par,   par2 = frank_cop$par2,   AIC = frank_cop$AIC),
  tibble(Copula = "Joe",          family = joe_cop$family,     par = joe_cop$par,     par2 = joe_cop$par2,     AIC = joe_cop$AIC)
) %>%
  mutate(
    dep_parameter = ifelse(is.na(par2), 
                           sprintf("%.4f", par),
                           sprintf("par=%.4f, par2=%.4f", par, par2))
  ) %>%
  select(Copula, dep_parameter, AIC) %>%
  arrange(AIC)

copula_tbl

# Select the Copula with the minimum AIC (in this scenario, the Joe copula)
X_con_copula <- joe_cop
```

# Scenario B Copula Fits and Comparisons

```{r}

u <- pull(Y_con_gpa, y_cdf)
v <- pull(Y_con_gpa, x_cdf)

gumbel_cop <- BiCopEst(u, v, family = 4)
print(gumbel_cop$AIC)
plot(gumbel_cop, type = "contour", margins = "unif", main = "Gumbel", col = "blue")
points(x=u, y=v, pch = 16)

t_cop <- BiCopEst(u, v, family = 2)
print(t_cop$AIC)
plot(t_cop, type = "contour", margins = "unif", main = "Student's t", col = "blue")
points(x=u, y=v, pch = 16)

gauss_cop <- BiCopEst(u, v, family = 1)
print(gauss_cop$AIC)
plot(gauss_cop, type = "contour", margins = "unif", main = "Gaussian", col = "blue")
points(x=u, y=v, pch = 16)

clayton_cop <- BiCopEst(u, v, family = 3)
print(clayton_cop$AIC)
plot(clayton_cop, type = "contour", margins = "unif", main = "Clayton", col = "blue")
points(x=u, y=v, pch = 16)

frank_cop <- BiCopEst(u, v, family = 5)
print(frank_cop$AIC)
plot(frank_cop, type = "contour", margins = "unif", main = "Frank", col = "blue")
points(x=u, y=v, pch = 16)

joe_cop <- BiCopEst(u, v, family = 6)
print(joe_cop$AIC)
plot(joe_cop, type = "contour", margins = "unif", main = "Joe", col = "blue")
points(x=u, y=v, pch = 16)

# Create summary table
copula_tbl <- dplyr::bind_rows(
  tibble(Copula = "Gumbel",       family = gumbel_cop$family,  par = gumbel_cop$par,  par2 = gumbel_cop$par2,  AIC = gumbel_cop$AIC),
  tibble(Copula = "Student's t",  family = t_cop$family,       par = t_cop$par,       par2 = t_cop$par2,       AIC = t_cop$AIC),
  tibble(Copula = "Gaussian",     family = gauss_cop$family,   par = gauss_cop$par,   par2 = gauss_cop$par2,   AIC = gauss_cop$AIC),
  tibble(Copula = "Clayton",      family = clayton_cop$family, par = clayton_cop$par, par2 = clayton_cop$par2, AIC = clayton_cop$AIC),
  tibble(Copula = "Frank",        family = frank_cop$family,   par = frank_cop$par,   par2 = frank_cop$par2,   AIC = frank_cop$AIC),
  tibble(Copula = "Joe",          family = joe_cop$family,     par = joe_cop$par,     par2 = joe_cop$par2,     AIC = joe_cop$AIC)
) %>%
  mutate(
    dep_parameter = ifelse(is.na(par2), 
                           sprintf("%.4f", par),
                           sprintf("par=%.4f, par2=%.4f", par, par2))
  ) %>%
  select(Copula, dep_parameter, AIC) %>%
  arrange(AIC)

copula_tbl

# Select the Copula with the minimum AIC (in this scenario, the Joe copula)
Y_con_copula <- joe_cop
```

# Contour Line Values for Plotting

This code block is for Neptune

```{r}
n <- 60
u <- v <- seq(0.02, 0.98, length.out = n)

# density on grid
z <- outer(u, v, function(a, b)
  BiCopPDF(a, b,
           family = marengo_cop$family,
           par    = marengo_cop$par,
           par2   = marengo_cop$par2)
)

lev <- c(0.2, 0.6, 1, 1.5, 2, 3, 5, 7)

contour(u, v, z, levels = lev, col = "blue",
        xlab = "u", ylab = "v",
        main = sprintf("BiCop density contours", marengo_cop$family))
lev

cls <- contourLines(u, v, z, levels = lev)

df <- do.call(rbind, lapply(seq_along(cls), function(i) {
  data.frame(
    level = cls[[i]]$level,
    x     = cls[[i]]$x,
    y     = cls[[i]]$y,
    line  = i,                        # optional: which polyline (useful if multiple per level)
    pt    = seq_along(cls[[i]]$x)     # optional: point index along the line
  )
}))

# If you truly want only 3 columns, drop line/pt:
# df <- df[, c("level","x","y")]

write.csv(df, "coastal_contours.csv", row.names = FALSE)
```

# Copula Simulation and Transformation

## Generate Samples

Generate a large number of samples). Split the number of samples across the two conditional copula models in proportion to the number of events in each dataset.

```{r}
N <- 10^6
n_A <- nrow(marshalltown_con) # Number of coincident events in scenario A
n_B <- nrow(marengo_con) # Number of coincident events in scenario B
n1 <- round((N * n_A) / (n_A + n_B), 0)
n2 <- N - n1

# Simulate U(0,1)^2 from each fitted copula
s1 <- VineCopula::BiCopSim(n1, obj = marshalltown_cop)
s2 <- VineCopula::BiCopSim(n2, obj = marengo_cop)

#plot(head(s1, 1000))
```

## Transform to Original Scale using Quantile Functions

Transform the simulated uniforms to original units using quantile functions.

```{r}
# Scenario A
marshalltown_gev <- lmomco::vec2par(marshalltown_gev, type = "gev")
marengo_glo <- lmomco::vec2par(marengo_glo, type = "glo")

marshalltown_A <- quagev(s1[,1], marshalltown_gev)
marengo_A <- quaglo(s1[,2], marengo_glo)

copula_sample_A <- data.frame(Marshalltown = marshalltown_A, Marengo = marengo_A)

# Scenario B
marengo_gev <- vec2par(marengo_gev, type = "gev")
marshalltown_glo <- vec2par(marshalltown_glo, type = "glo")

marengo_B <- quagev(s2[,1], marengo_gev)
marshalltown_B <- quaglo(s2[,2], marshalltown_glo)

copula_sample_B <- data.frame(Marshalltown = marshalltown_B, Marengo = marengo_B)

# Combine samples from both scenarios; used in KDE 
copula_sample <- rbind(copula_sample_A, copula_sample_B)
```

# Isoline Derivation

## Generate Unit Square

```{r}
# Using "low" resolution here to reduce run times
x <- c(10^(-5), seq(999.9*10^(-5), 1-(1*10^(-6)), 10^(-4)))
y <- c(10^(-5), seq(999.9*10^(-5), 1-(1*10^(-6)), 10^(-4)))
grid <- list(x = x, y = y)
u <- expand.grid(x, y)
```

## Evaluate Copula CDF

Evaluate the copula CDF at each point on the grid.

```{r}
# Evaluates CDF of the specified parametric copula
uCopA <- VineCopula::BiCopCDF(u[,1], u[,2], marshalltown_cop) 
uCopB <- VineCopula::BiCopCDF(u[,1], u[,2], marengo_cop) 
```

## Determine Return Period

Determine the return period of each point on the unit square.

This code block defines the sampling frequency for daily data and a function to compute the return period surface matrix for an x, y grid in column blocks.

```{r}
mu <- 24 * 365.25 # Sampling frequency for hourly data

# Computes return period surface matrix z on a grid of x and y values in blocks of columns
make_z_surface_blocked <- function(x, y, Cmat, mean_waiting_time, block = 500L){
  nx <- length(x); ny <- length(y)
  z <- matrix(NA_real_, nrow = nx, ncol = ny) # Pre-allocate output matrix z

  # Iterate over the y-axis in chunks
  for (j0 in seq.int(1L, ny, by = block)) {
    j1 <- min(j0 + block - 1L, ny)
    yb <- y[j0:j1]

    # denom is nx × nb (nb = block size)
    denom <- 1 - x - rep(yb, each = nx) + as.vector(Cmat[, j0:j1, drop=FALSE])
    z[, j0:j1] <- mean_waiting_time / matrix(denom, nrow = nx, ncol = length(yb))
  }
  z
}
```

### Scenario A Return Period

```{r}

# Mean waiting time is the average time, in days, between events
mean_waiting_time_A <- mu / rate_marshalltown 

# Reshape into a matrix: rows correspond to x, columns correspond to y
# IMPORTANT: expand.grid(x,y) varies x fastest, then y
Cmat <- matrix(uCopA, nrow = length(x), ncol = length(y), byrow = FALSE)

z  <- make_z_surface_blocked(x, y, Cmat, mean_waiting_time_A, block=200L)

rp_50 <- 50 # return period in years
xy50 <- grDevices::contourLines(x, y, z, levels = mu * rp_50)

if (length(xy50) == 0) stop("No contour found for RP = ", rp_50)

contourLines_A_50 <- do.call(rbind, lapply(seq_along(xy50), function(i){
  data.frame(id=i, level=xy50[[i]]$level, marshalltown_cdf=xy50[[i]]$x, marengo_cdf=xy50[[i]]$y, rp = rp_50)
}))


rp_100 <- 100
xy100 <- grDevices::contourLines(x, y, z, levels = mu * rp_100)

contourLines_A_100 <- do.call(rbind, lapply(seq_along(xy100), function(i){
  data.frame(id=i, level=xy100[[i]]$level, marshalltown_cdf=xy100[[i]]$x, marengo_cdf=xy100[[i]]$y, rp = rp_100)
}))
```

### Scenario B Return Period

```{r}

# Mean waiting time is the average time, in days, between events
mean_waiting_time_B <- mu / rate_marengo

# Reshape into a matrix: rows correspond to x, columns correspond to y
# IMPORTANT: expand.grid(x,y) varies x fastest, then y
Cmat <- matrix(uCopB, nrow = length(x), ncol = length(y), byrow = FALSE)

z  <- make_z_surface_blocked(x, y, Cmat, mean_waiting_time_B, block=200L)

rp_50 <- 50 # return period in years
xy50 <- grDevices::contourLines(x, y, z, levels = mu * rp_50)

if (length(xy50) == 0) stop("No contour found for RP = ", rp_50)

contourLines_B_50 <- do.call(rbind, lapply(seq_along(xy50), function(i){
  data.frame(id=i, level=xy50[[i]]$level, marshalltown_cdf=xy50[[i]]$x, marengo_cdf=xy50[[i]]$y, rp = rp_50)
}))


rp_100 <- 100
xy100 <- grDevices::contourLines(x, y, z, levels = mu * rp_100)

contourLines_B_100 <- do.call(rbind, lapply(seq_along(xy100), function(i){
  data.frame(id=i, level=xy100[[i]]$level, marshalltown_cdf=xy100[[i]]$x, marengo_cdf=xy100[[i]]$y, rp = rp_100)
}))
```

## Transform Variates to Original Scale

### Scenario A Transform Variates

```{r}

contourLines_A <- rbind(contourLines_A_50, contourLines_A_100)
p_pds <- contourLines_A$marshalltown_cdf
p_marshalltown_A <- langbein_adjustment(rate_marshalltown, p_pds) # Annualized probabilities for Marshalltown

contourLines_A <- contourLines_A %>%
  mutate(marshalltown_annualized_p = p_marshalltown_A,
         marshalltown = quagev(marshalltown_annualized_p, marshalltown_gev),
         marengo = quaglo(marengo_cdf, marengo_glo),
         scenario = "A")
```

### Scenario B Transform Variates

```{r}

contourLines_B <- rbind(contourLines_B_50, contourLines_B_100)
p_pds <- contourLines_B$marengo_cdf
p_marengo_B <- langbein_adjustment(rate_marengo, p_pds) # Annualized probabilities for Marengo

contourLines_B <- contourLines_B %>%
  mutate(marengo_annualized_p = p_marengo_B,
         marengo = quagev(marengo_annualized_p, marengo_gev),
         marshalltown = quaglo(marshalltown_cdf, marshalltown_glo),
         scenario = "B")
```

## Plot Isolines

```{r}

ggplot(contourLines_A, aes(marshalltown, marengo, group = rp)) +
  geom_path(linewidth = 1) +
  theme_bw() +
  xlab("Marshalltown Flow (cfs)") +
  ylab("Marengo Flow (cfs)") +
  ggtitle("Scenario A: Extremal Marshalltown Flows") +
  xlim(0, 40000) +
  ylim(0, 40000) +
  coord_fixed(ratio = 1)

ggplot(contourLines_B, aes(marengo, marshalltown, group = rp)) +
  geom_path(linewidth = 1) +
  theme_bw() +
  xlab("Marengo Flow (cfs)") +
  ylab("Marshalltown Flow (cfs)") +
  ggtitle("Scenario B: Extremal Marengo Flows") +
  xlim(0, 50000) +
  ylim(0, 50000) +
  coord_fixed(ratio = 1)

contourLines_A_common_cols <- contourLines_A %>%
  select(rp, marshalltown, marengo, scenario)

contourLines_B_common_cols <- contourLines_B %>%
  select(rp, marshalltown, marengo, scenario)

contourLines <- rbind(contourLines_A_common_cols, contourLines_B_common_cols)

ggplot(contourLines, aes(marshalltown, marengo, color = scenario, linetype = factor(rp))) +
  geom_path(linewidth = 1) +
    scale_color_manual(values = c(
    "A" = "darkgreen",
    "B" = "purple")) +
  scale_linetype_manual(values = c(
    "50" = "dashed",
    "100" = "solid")) +
  xlim(0, 50000) +
  ylim(0, 50000) +
  theme_bw() +
  labs(color = "Scenario", linetype = "Return Period (years)") +
  xlab("Marshalltown Flow (cfs)") +
  ylab("Marengo Flow (cfs)") +
  ggtitle("") +
  coord_fixed(ratio = 1)

# View 1/100 joint probability contour from scenario A
# contourLines[contourLines$scenario == "A" & contourLines$rp == "100", ]
```

# Compute Envelope Curve

## Combine Isolines into Envelope Curve

The make_envelope function takes a data frame containing isoline points for one return period, rp, but multiple curves, distinguished by scenario, and produces a new curve that is the upper envelope (for each x-value, it keeps the maximum y-value across all scenarios).

```{r}

make_envelope <- function(df_rp, n_grid = 500L) {
  
  # Clean and sort input 
  # Drop rows where marshalltown (x) or marengo (y) are missing/non-finite
  # Sorts by marshalltown values
  df_rp <- df_rp %>%
    filter(is.finite(marshalltown), is.finite(marengo)) %>%
    arrange(marshalltown)

  # Create a common x-grid over the range covered by the isolines
  # Builds xg, a sequence from the minimum to maximum marshalltown in the data,        with n_grid points
  xg <- seq(min(df_rp$marshalltown), max(df_rp$marshalltown), length.out = n_grid)

  # Interpolate each scenario separately (safer than interpolating across mixed        curves)
  # Splits data by scenario
  # For each scenario:
    # Sorts by marshalltown
    # Removes duplicate x-values so interpolation behaves well
    # Use approx() to linearly interpolate marengo values at every xg
    # rule = 1 means no extrapolation (any xout < min(x) or xout > max(x) returns      NA)
  y_by_curve <- df_rp %>%
    group_by(scenario) %>%
    group_split() %>%
    map(function(d) {
      d <- d %>% arrange(marshalltown)
      # approx requires unique x to behave well
      d <- d %>% distinct(marshalltown, .keep_all = TRUE)

      approx(x = d$marshalltown, y = d$marengo, xout = xg,
             rule = 1, ties = "ordered")$y
    })

  # Compute the envelope = pointwise maximum across curves
  yenv <- do.call(pmax, c(y_by_curve, list(na.rm = TRUE)))

  # Return a table
  tibble(
    rp = unique(df_rp$rp),
    marshalltown = xg,
    marengo = yenv) %>%
    dplyr::filter(marshalltown >= 0, marengo >= 0)
}
```

Apply make_envelope function to combine isolines in contourLines into an envelope curve.

```{r}

envelope_curves <- contourLines %>%
  group_by(rp) %>%
  group_split() %>%
  map_dfr(make_envelope)

envelope_curves
```

## Plot Individual Contour Lines and Envelope Curve

```{r}

contourLinesPlot <- ggplot() +
  geom_path(
    data = contourLines,
    aes(marshalltown, marengo, color = scenario, linetype = factor(rp)),
    linewidth = 0.8, alpha = 0.7
  ) +
  geom_path(
    data = envelope_curves,
    aes(marshalltown, marengo, linetype = factor(rp)),
    color = "black", linewidth = 1.2
  ) +
  scale_color_manual(values = c(
    "A" = "darkgreen",
    "B" = "purple"
  )) +
  scale_linetype_manual(values = c(
    "50"  = "dashed",
    "100" = "solid"
  )) +
  theme_bw() +
  coord_fixed(ratio = 1) +
  labs(
    color = "Scenario",
    linetype = "Return Period (years)",
    x = "Marshalltown Flow (cfs)",
    y = "Marengo Flow (cfs)"
  )

contourLinesPlot
```

# Compute Relative Density Along Isolines

## KDE Along Isoline

Compute the KDE values along an isoline in *flow space* using your objects:

-   Iso: data.frame/tibble with columns marshalltown, marengo (one RP + one scenario, or already merged)

-   cop_sample: your simulated samples data.frame with columns Marshalltown, Marengo (flow space)

-   Data: optional observed data in flow space with same two columns; used when Source="Data"

-   Source: "Sample" (default; uses copula simulations) or "Data" (uses observed pairs)

-   Sim_Max: truncation multiplier (same intent as your original function)

    -   With very large simulations, heavy-tailed marginals can generate rare but enormous values. Sim_Max applies a guardrail by keeping the KDE focused on the range relevant to your contour/isoline or observed data.

```{r}

kde_along_isoline <- function(Iso,
                              cop_sample,
                              Data = NULL,
                              Source = c("Sample", "Data"),
                              Sim_Max = 10,
                              gridsize = c(200, 200)) {

  Source <- match.arg(Source)

  # Keep exactly these rows and order
  Iso2 <- Iso %>% dplyr::select(marshalltown, marengo)

  # eval.points must be an n×2 numeric matrix for ks::kde
  Iso_mat <- as.matrix(Iso2)
  storage.mode(Iso_mat) <- "double"

  if (Source == "Sample") {
    xdat <- cop_sample %>% dplyr::select(Marshalltown, Marengo) %>% as.matrix()
    storage.mode(xdat) <- "double"

    ref <- if (!is.null(Data)) Data %>% dplyr::select(Marshalltown, Marengo) else Iso2

    remove <- which(
      xdat[,1] > Sim_Max * max(ref[[1]], na.rm=TRUE) |
      xdat[,2] > Sim_Max * max(ref[[2]], na.rm=TRUE)
    )
    if (length(remove) > 0) xdat <- xdat[-remove, , drop = FALSE]
  } else {
    if (is.null(Data)) stop("When Source='Data', supply `Data`.")
    xdat <- Data %>% dplyr::select(Marshalltown, Marengo) %>% na.omit() %>% as.matrix()
    storage.mode(xdat) <- "double"
  }

  est <- ks::kde(x = xdat, eval.points = Iso_mat, gridsize = gridsize)$estimate

  if (length(est) != nrow(Iso_mat)) {
    stop("KDE length mismatch: got ", length(est), " but Iso has ", nrow(Iso_mat), " rows.")
  }

  rng <- range(est, na.rm = TRUE)
  scaled <- if (is.finite(rng[1]) && is.finite(rng[2]) && rng[2] > rng[1]) {
    (est - rng[1]) / (rng[2] - rng[1])
  } else rep(NA_real_, length(est))

  list(prediction = est, contour = scaled)
}
```

## Find Design Event Along Each Isoline

Fine "design events" or "most likely events" on an isoline. The "most likely event" is the isoline point with the maximum KDE.

-   Iso: data.frame with marshalltown, marengo columns

-   prediction: KDE values returned by kde_along_isoline()\$prediction

-   N_Ensemble: number of sampled ensemble points along isoline (weighted by prediction)

```{r}

extract_design_events_2d <- function(Iso, prediction, N_Ensemble = 0) {

  Iso_mat <- Iso %>%
    dplyr::select(marshalltown, marengo) %>%
    as.data.frame()

  # Most likely = max KDE point on isoline
  idx <- which(prediction == max(prediction, na.rm = TRUE))[1]
  most <- data.frame(
    marshalltown = Iso_mat[idx, 1],
    marengo      = Iso_mat[idx, 2]
  )

  # Keep your previous "full dependence" behavior (max x and max y on isoline)
  full <- data.frame(
    marshalltown = max(Iso_mat[, 1], na.rm = TRUE),
    marengo      = max(Iso_mat[, 2], na.rm = TRUE)
  )

  ens <- NULL # Initialize ensemble output as NULL
  if (N_Ensemble > 0) {
    # Find indices of isoline points where the KDE values are finite and strictly positive
    ok <- which(is.finite(prediction) & prediction > 0)
    # If there are no valid KDE weights, return an empty 0-row matrix/data frame with the same number of columns as Iso_mat
    if (length(ok) == 0) {
      ens <- Iso_mat[0, , drop = FALSE]
    # Randomly select N_Ensemble indices from the valid set ok
    } else {
      # prob = prediction[ok] makes it weighted sampling: higher KDE density -> more likely to be chosen
      draw <- sample(ok, size = N_Ensemble, replace = TRUE, prob = prediction[ok])
      # Extracts the sampled isoline coordinates corresponding to the chosen indices
      ens <- Iso_mat[draw, , drop = FALSE]
      # Label the two columns
      names(ens) <- c("marshalltown", "marengo")
    }
  }

  list(MostLikelyEvent = most, FullDependence = full, Ensemble = ens)
}
```

## KDE and Design Event for 1/100 AEP Contour

```{r}
Iso100 <- envelope_curves %>% dplyr::filter(rp == 100)   # or filter(scenario=="A", rp==100)

kde_out100 <- kde_along_isoline(
  Iso = Iso100,
  cop_sample = copula_sample,     # columns: Marshalltown, Marengo
  gridsize = c(400, 400)
)

design100 <- extract_design_events_2d(Iso100, kde_out100$prediction, N_Ensemble = 200)
designEvent100 <- design100$MostLikelyEvent

contourLines_kde_100 <- envelope_curves %>%
  dplyr::group_by(rp) %>%
  dplyr::group_split() %>%
  purrr::map_dfr(function(df_iso) {

    kde_out_i <- kde_along_isoline(
      Iso        = df_iso,
      cop_sample = copula_sample,
      Source     = "Sample",
      Sim_Max    = 10,
      gridsize   = c(400, 400)
    )

    stopifnot(length(kde_out_i$prediction) == nrow(df_iso))

    df_iso %>%
      dplyr::mutate(
        kde        = kde_out_i$prediction,
        kde_scaled = kde_out_i$contour
      )
  })
```

## KDE and Design Event for 1/50 AEP Contour

```{r}
Iso50 <- envelope_curves %>% dplyr::filter(rp == 50)   

kde_out50 <- kde_along_isoline(
  Iso = Iso50,
  cop_sample = copula_sample,     # columns: Marshalltown, Marengo
  gridsize = c(400, 400)
)

design50 <- extract_design_events_2d(Iso50, kde_out50$prediction, N_Ensemble = 200)
designEvent50 <- design50$MostLikelyEvent

contourLines_kde_50 <- envelope_curves %>%
  dplyr::group_by(rp) %>%
  dplyr::group_split() %>%
  purrr::map_dfr(function(df_iso) {

    kde_out_i <- kde_along_isoline(
      Iso        = df_iso,
      cop_sample = copula_sample,
      Source     = "Sample",
      Sim_Max    = 10,
      gridsize   = c(400, 400)
    )

    stopifnot(length(kde_out_i$prediction) == nrow(df_iso))

    df_iso %>%
      dplyr::mutate(
        kde        = kde_out_i$prediction,
        kde_scaled = kde_out_i$contour
      )
  })
```

# Final Plot

Plot the envelope isolines colored by relative density (KDE-scaled), with return period labels along the curve.

-   Overlays:

    -   a subset of simulated pairs (grey),

    -   observed coincident-event pairs from both scenarios (different shapes/colors),

    -   the “most likely event” marker on the isoline.

```{r}

marshalltown_con_data <- marshalltown_con %>%
  mutate(Scenario = "Marshalltown_Conditioning")

marengo_con_data <- marengo_con %>%
  mutate(Scenario = "Marengo_Conditioning")

data <- rbind(marshalltown_con_data, marengo_con_data) %>%
  select(Marshalltown, Marengo, Scenario)

contourLines_kde_100 <- contourLines_kde_100 %>%
  mutate(aep = 1/rp)

ggplot(contourLines_kde_100, aes(marshalltown, marengo)) +
    geom_point(
    data = head(copula_sample, n = 10000),
    aes(x = Marshalltown, y = Marengo),
    inherit.aes = FALSE,
    color = "darkgrey", shape = 21
  ) +
  geom_path(aes(group = aep, color = kde_scaled), size = 1.2) +
  scale_color_distiller(palette = "RdYlBu", direction = -1,
                        name = "Relative Probability Density", limits = c(0, 1)) +
  geom_point(
    data = data,
    aes(x = Marshalltown, y = Marengo, shape = Scenario),
    inherit.aes = FALSE,
    color = c(Marshalltown_Conditioning = "darkgreen", Marengo_Conditioning = "purple")[data$Scenario],
    size = 2, stroke = 1
  ) +
  scale_shape_manual(values = c(Marshalltown_Conditioning = 1, Marengo_Conditioning = 4), name = "Scenario") +
geom_point(
    data = design100$MostLikelyEvent,
    aes(x = marshalltown, y = marengo, fill = "Most Likely Event"),
    inherit.aes = FALSE,
    color = "black",
    shape = 23,
    linewidth = 2,
    stroke = 1.1
  ) +
  geom_point(
    data = design50$MostLikelyEvent,
    aes(x = marshalltown, y = marengo, fill = "Most Likely Event"),
    inherit.aes = FALSE,
    color = "black",
    shape = 23,
    linewidth = 2,
    stroke = 1.1
  ) +
    geomtextpath::geom_textpath(
    aes(label = aep, group = rp, color = kde_scaled),
    color = "black", # label color
    linetype = 0,   # (or) linetype = 0
    linewidth = 4,
    text_smoothing = 10,     # helps spacing along the curve
    vjust = -0.2
  ) +
  scale_fill_manual(
    values = c("Most Likely Event" = "red"),
    name = ""
  ) +
    theme_bw() + coord_fixed(1) +
  xlim(0, 80000) + ylim(0, 80000) +
  labs(x = "Marshalltown Flow (cfs)", y = "Marengo Flow (cfs)") 
```
