---
title: "ordinal_association"
author: "Avital Breverman"
format: html
editor: visual
---

```{r}
rm(list = ls())

library(dssrip2)
library(MultiHazard)
library(dplyr)

options(java.parameters = "-Xmx16000M")
dss_connect("C:/Programs/hec-monolith", monolith = TRUE)
```

# Purpose

Determine ordinal association as computed using Kendall's tau between two variables (flow and precipitation) over a range of lags. The lag that produces the highest Kendall's tau is used to identify events that are coincident with the extremal events.

# Import Data

```{r}
setwd("C:/Projects/Neptune/CopulaAnalysis/Examples/TygartPhillippiInteriorDrainage")
dss_file_path <- "C:/Projects/Neptune/CopulaAnalysis/Examples/TygartPhillippiInteriorDrainage/Phillippi/Phillippi.dss"

dss_paths <- dssrip2::dss_catalog(dss_file_path)
dss_paths

# Rolling 3-day accumulation of precipitation (USC00466982)
precipitation_3day_accumulation <- dss_read(dss_file_path, dss_paths[10])

# Annual max 3-day precipitation
precipitation_3day_ams <- dss_read(dss_file_path, dss_paths[9])

# Daily flow (USGS 03054500)
flow_daily <- dss_read(dss_file_path, dss_paths[1])

# Annual max daily flow
flow_daily_ams <- dss_read(dss_file_path, dss_paths[2])
```

# Create Dataframe

```{r}

continuous_time_series <- left_join(flow_daily, precipitation_3day_accumulation, by = "datetime") %>%
  na.omit()
```

# Rank Association

Use the MultiHazard package's Kendall_Lag function to compute Kendall's tau and p-values over a range of lags (-4 to +4 days, incremented by 1 day).

```{r}
lag <- 4 # days

# Calculate Kendall's tau at different lags
kendall_tau_lag_results <- MultiHazard::Kendall_Lag(Data = continuous_time_series,
                      Lags = seq(-lag, lag, 1),  
                      PLOT = TRUE, # display plot
                      GAP = 0.1) # length of y-axis above and below max and min                                       

tau_values <- kendall_tau_lag_results$Value
p_values <- kendall_tau_lag_results$Test

# View tau coefficients
print(tau_values)
print(p_values)
```

# Apply the Lag

```{r}

continuous_time_series <- continuous_time_series %>%
  mutate(precipitation_lag_plus1day = lead(precipitation, 1),
         flow_lag_minus1day = lag(flow, 1))

flow_dominant <- continuous_time_series %>%
  mutate(
    datetime = as.Date(datetime),
    water_year = if_else(
      format(datetime, "%m-%d") >= "10-01",
      as.integer(format(datetime, "%Y")) + 1L,
      as.integer(format(datetime, "%Y"))
    )
  ) %>%
  group_by(water_year) %>%
  slice_max(order_by = flow, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(water_year, datetime, flow, precipitation_lag_plus1day)


precipitation_dominant <- continuous_time_series %>%
  mutate(
    datetime = as.Date(datetime),
    water_year = if_else(
      format(datetime, "%m-%d") >= "10-01",
      as.integer(format(datetime, "%Y")) + 1L,
      as.integer(format(datetime, "%Y"))
    )
  ) %>%
  group_by(water_year) %>%
  slice_max(order_by = precipitation, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(water_year, datetime, precipitation, flow_lag_minus1day)

```

# Write Datasets to csv

```{r}
write.csv(flow_dominant, file = "flow_dominant.csv")
write.csv(precipitation_dominant, file = "precipitation_dominant.csv")
```
