---
title: "ordinal_association"
author: "Avital Breverman"
format: html
editor: visual
---

```{r}
rm(list = ls())

library(dssrip2)
library(MultiHazard)
library(dplyr)
library(ggplot2)

options(java.parameters = "-Xmx16000M")
dss_connect("C:/Programs/hec-monolith", monolith = TRUE)
```

# Purpose

In this example, the two variables are river flow and precipitation. We will quantify the ordinal association between extreme events in one variable and near-coincident events in the other using Kendall’s tau. The analysis will be performed in both directions to address the following questions:

-   What time lag between an extreme precipitation event and a high-flow event produces the strongest association?

-   What time lag between an extreme river-flow event and a large precipitation event produces the strongest association?

For each direction, we will evaluate a range of positive and negative time lags. The lag associated with the highest Kendall’s tau will be interpreted as the temporal offset at which the extreme events are most strongly associated.

# Import Data

```{r}
setwd("C:/Projects/Neptune/CopulaAnalysis/Examples/")
dss_file_path <- "C:/Projects/Neptune/CopulaAnalysis/Examples/Philippi_Interior_Drainage/Philippi_Interior_Drainage.dss"

dss_paths <- dssrip2::dss_catalog(dss_file_path)
dss_paths

# Rolling 3-day accumulation of precipitation (USC00466982)
precipitation_3day_accumulation <- dss_read(dss_file_path, dss_paths[6])

# Annual max 3-day precipitation
precipitation_3day_ams <- dss_read(dss_file_path, dss_paths[4])

# Daily flow (USGS 03054500)
flow_daily <- dss_read(dss_file_path, dss_paths[8])

# Annual max daily flow
flow_daily_ams <- dss_read(dss_file_path, dss_paths[9])
```

# Create Dataframes

```{r}

precipitation_extremal <- left_join(precipitation_3day_ams, flow_daily, by = "datetime") %>% na.omit()

flow_extremal <- left_join(flow_daily_ams, precipitation_3day_accumulation, by = "datetime") %>% na.omit()
```

# Rank Association for Precipitation Extremal

Use the MultiHazard package's Kendall_Lag function to compute Kendall's tau and p-values over a range of lags (-5 to +5 days, incremented by 1 day).

```{r}
max_lag <- 5 # days

# Calculate Kendall's tau at different lags
kendall_tau_lag_results <- MultiHazard::Kendall_Lag(Data = precipitation_extremal,
                      Lags = seq(-max_lag, max_lag, 1),  
                      PLOT = TRUE, # display plot
                      GAP = 0.1) # length of y-axis above and below max and min                                       

tau_values <- kendall_tau_lag_results$Value
p_values <- kendall_tau_lag_results$Test

# View tau coefficients
print(tau_values)
print(p_values)
```

# Rank Association for Flow Extremal

```{r}

# Calculate Kendall's tau at different lags
kendall_tau_lag_results <- MultiHazard::Kendall_Lag(Data = flow_extremal,
                      Lags = seq(-max_lag, max_lag, 1),  
                      PLOT = TRUE, # display plot
                      GAP = 0.1) # length of y-axis above and below max and min                                       

tau_values <- kendall_tau_lag_results$Value
p_values <- kendall_tau_lag_results$Test

# View tau coefficients
print(tau_values)
print(p_values)
```

# Apply the Lag

Flows with the highest correlation to extremal precipitation occur one day after as the extremal precipitation.

Precipitation events with the highest correlation to extremal flows occur on the same day as the extremal flows.

```{r}
flow_extremal_lag_precip <- 0 # days
precip_extremal_lag_flow <- 1 # days

# Prepare date-only daily time series for reliable joins.
precipitation_daily <- precipitation_3day_accumulation %>%
  transmute(
    date = as.Date(datetime),
    precipitation = precipitation
  ) %>%
  distinct(date, .keep_all = TRUE)

flow_daily_values <- flow_daily %>%
  transmute(
    date = as.Date(datetime),
    flow = flow
  ) %>%
  distinct(date, .keep_all = TRUE)

# Prepare date-only annual-maxima event tables.
flow_extremes <- flow_daily_ams %>%
  transmute(
    flow_event_date = as.Date(datetime),
    flow_extreme = flow
  ) %>%
  distinct(flow_event_date, .keep_all = TRUE)

precipitation_extremes <- precipitation_3day_ams %>%
  transmute(
    precipitation_event_date = as.Date(datetime),
    precipitation_extreme = precipitation
  ) %>%
  distinct(precipitation_event_date, .keep_all = TRUE)

```

### **Pair each extremal flow with same-day precipitation**

```{r}
flow_extremal_pairs <- flow_extremes %>%
  mutate(
    precipitation_date = flow_event_date + flow_extremal_lag_precip
  ) %>%
  left_join(
    precipitation_daily,
    by = c("precipitation_date" = "date")
  ) %>%
  filter(!is.na(precipitation)) %>%
  transmute(
    precipitation_date,
    flow_date = flow_event_date,
    precipitation,
    flow = flow_extreme,
    event_type = "Extremal flow; same-day precipitation"
  )
```

### **Pair each extremal precipitation event with next-day flow**

```{r}
precipitation_extremal_pairs <- precipitation_extremes %>%
  mutate(
    flow_date = precipitation_event_date + precip_extremal_lag_flow
  ) %>%
  left_join(
    flow_daily_values,
    by = c("flow_date" = "date")
  ) %>%
  filter(!is.na(flow)) %>%
  transmute(
    precipitation_date = precipitation_event_date,
    flow_date,
    precipitation = precipitation_extreme,
    flow,
    event_type = "Extremal precipitation; next-day flow"
  )
```

```{r}
event_pairs_plot <- bind_rows(
  flow_extremal_pairs,
  precipitation_extremal_pairs
)

ggplot(
  event_pairs_plot,
  aes(
    x = precipitation,
    y = flow,
    color = event_type,
    shape = event_type
  )
) +
  geom_point(size = 3, alpha = 0.8) +
  scale_color_manual(
    values = c(
      "Extremal flow; same-day precipitation" = "blue",
      "Extremal precipitation; next-day flow" = "red"
    )
  ) +
  scale_shape_manual(
    values = c(
      "Extremal flow; same-day precipitation" = 16,
      "Extremal precipitation; next-day flow" = 15
    )
  ) +
  labs(
    title = "Precipitation–Flow Event Pairs",
    subtitle = paste0(
      "Extremal-flow to precipitation lag: ", flow_extremal_lag_precip,
      " days; extremal-precipitation to flow lag: ",
      precip_extremal_lag_flow, " day"
    ),
    x = "3-Day Precipitation",
    y = "Daily Flow",
    color = "Event Pair",
    shape = "Event Pair"
  ) +
  theme_minimal()
```

# Write Datasets to csv

```{r}
write.csv(flow_extremal_pairs, file = "flow_dominant.csv")
write.csv(precipitation_extremal_pairs, file = "precipitation_dominant.csv")
```

# 
