---
title: "Copula Example"
format: html
editor: visual
author: Avital Breverman
---

```{r}
rm(list = ls())

library(dataRetrieval)
library(tidyverse)
library(POT)
library(lmom)
library(lmomco)
library(VineCopula)
library(MultiHazard)

# User directory 
dir <- "C:/Projects/Neptune/CopulaAnalysis/R"
setwd(dir)
```

-   Download daily streamflow from two gages on the mainstem of the Iowa River

-   Extract a partial duration series at each gage

-   Compute rank correlation coefficient (Kendall's tau) between the two locations

-   Define copula scenarios

-   Fit copula models and select copula for each scenario based on Akaike Information Criterion (AIC)

-   Generate samples from each copula and transform to original variable scale

-   Derive isoline for specified probabilities (return periods)

-   Combine isolines into envelope curve

-   Compute relative density of pairs on each isoline using kernel density estimation

# Functions

Define custom functions.

```{r}
# Computes a time separation criterion, in days, between events in a PDS to ensure independence. Equation presented in Bulletin 17.
# DA in sq mi
B17_days_between <- function(DA) {
  ceiling(5 + log(DA))
}

# Madsen annualization of Generalized Pareto distribution to Generalized Extreme Value distribution # Reference: Madsen, H., Rasmussen, P., & Rosbjerg, D. (1997). Comparison of annual maximum series and partial duration series methods for modeling extreme hydrologic events 1. At-site modeling. Water Resources Research, # 33(4), 747-757.
madsen_annualization <- function(gpd_par, lambda) {
  gev_xi <- gpd_par["xi"] + (gpd_par["alpha"] / gpd_par["k"]) * (1 - lambda ^ -gpd_par["k"]) # scale
  gev_alpha <- gpd_par["alpha"] * lambda ^ -gpd_par["k"] # scale
  gev_k <- gpd_par["k"]  
  
  print(gev_xi)
  print(gev_alpha)
  print(gev_k)
  
  c(xi = as.numeric(gev_xi),
  alpha = as.numeric(gev_alpha),
  k = as.numeric(gev_k))
}
```

# Download Streamflow Data using dataRetrieval

Download daily streamflow data for:

-   Iowa River at Marshalltown, IA (05451500)

-   Iowa River at Marengo, IA (05453100)

NWIS is scheduled for decommission but I'm rolling with these functions for as long as I can.

```{r}
marshalltown <- "05451500"
marengo <- "05453100"

marshalltown_daily <- readNWISdv(marshalltown, "00060") %>% renameNWISColumns() %>%
  na.omit()
marengo_daily <- readNWISdv(marengo, "00060") %>% renameNWISColumns() %>%
  na.omit()

daily_flow_joined <- left_join(marshalltown_daily, marengo_daily, by = "Date") %>%
  select(Date, Flow.x, Flow.y) %>%
  rename(Marshalltown = Flow.x,
         Marengo = Flow.y) %>%
  na.omit()
```

# Plot Daily Streamflow

```{r}
marshalltown_plot <- ggplot(marshalltown_daily, aes(x = Date, y = Flow)) +
  geom_line() +
  theme_bw() +
  xlab("") +
  ylab("Flow (cfs)") +
  ggtitle("Marshalltown")
marshalltown_plot

marengo_plot <- ggplot(marengo_daily, aes(x = Date, y = Flow)) +
  geom_line() +
  theme_bw() +
  xlab("") +
  ylab("Flow (cfs)") +
  ggtitle("Marengo")
marengo_plot
```

# Extract Marengo PDS

```{r}
# Create data frame with obs and time columns
pot_marengo <- data.frame(
  obs = marengo_daily$Flow,
  time = as.numeric(marengo_daily$Date)  # Convert to numeric
)

da_marengo <- 2794 # squuare miles
time_sep <- B17_days_between(da_marengo)
thresh_marengo <- 10000 # cfs

# Extract partial duration series
pds_marengo <- POT::clust(data = pot_marengo,
                              u = thresh_marengo, # threshold
                              tim.cond = time_sep, # time separation
                              clust.max = TRUE) # extract cluster maxima

summary(pds_marengo)
nrow(pds_marengo)

# Build dataframe
df_pds_marengo <- data.frame(pds_marengo) %>%
  rename(Date = time,
         Flow = obs) %>%
  mutate(Date = as.Date(Date))  %>%
  select(Date, Flow)

# Compute rate of events
years_marengo <-
  as.numeric((
    max(marengo_daily$Date) - min(marengo_daily$Date) + 1
  ) / 365.25)
rate_marengo <- nrow(df_pds_marengo) / years_marengo

# Plot daily flow and PDS
marengo_plot +
  geom_hline(yintercept = thresh_marengo, linetype = "dashed") +
  geom_point(data = df_pds_marengo, aes(x = Date, y = Flow),colour = "#A60F2D", shape = 19, size = 2) 
```

# Extract Marshalltown PDS

```{r}
# Create data frame with obs and time columns
pot_marshalltown <- data.frame(
  obs = marshalltown_daily$Flow,
  time = as.numeric(marshalltown_daily$Date)  # Convert to numeric
)

da_marshalltown <- 1532 # square miles
time_sep <- B17_days_between(da_marengo)
thresh_marshalltown <- 5500 # cfs

# Extract partial duration series
pds_marshalltown <- POT::clust(data = pot_marshalltown,
                              u = thresh_marshalltown, # threshold
                              tim.cond = time_sep, # time separation
                              clust.max = TRUE) # extract cluster maxima

summary(pds_marshalltown)
nrow(pds_marshalltown)

# Build dataframe
df_pds_marshalltown <- data.frame(pds_marshalltown) %>%
  rename(Date = time,
         Flow = obs) %>%
  mutate(Date = as.Date(Date)) %>%
  select(Date, Flow)

# Compute rate of events
years_marshalltown <-
  as.numeric((
    max(marshalltown_daily$Date) - min(marshalltown_daily$Date) + 1
  ) / 365.25)
rate_marshalltown <- nrow(df_pds_marshalltown) / years_marshalltown

# Plot daily flow and PDS
marshalltown_plot +
  geom_hline(yintercept = thresh_marshalltown, linetype = "dashed") +
  geom_point(data = df_pds_marshalltown, aes(x = Date, y = Flow),colour = "#A60F2D", shape = 19, size = 2) 
```

# Rank Correlation: Kendall's tau

Compute rank correlation between flows at Marshalltown and Marengo using Kendall's rank correlation coefficient (Kendall's tau) over a range of lags.

I computed Kendall's tau under 2 scenarios:

-   Rank correlation between the PDS of Marengo flows and lagged daily flows at Marshalltown

-   Rank correlation between the Marshalltown flows and lagged daily flows at Marengo

In both scenarios, the Kendall's tau value was highest at a lag of 0 (flows on the same date).

```{r}

lag <- 5 # -5 to +5 day lags

# Prepare data frame with 3 columns: Date, Variable1, Variable2
# For auto-correlation, use same variable twice
marengo_con <- left_join(df_pds_marengo, marshalltown_daily, by = "Date") %>%
  select(Date, Flow.x, Flow.y) %>%
  rename(Marengo = Flow.x,
         Marshalltown = Flow.y) %>%
  na.omit()

head(marengo_con)

# Calculate Kendall's tau at different lags
result <- MultiHazard::Kendall_Lag(Data = marengo_con,
                      Lags = seq(-lag, lag, 1),  # lag is applied to second named variable
                      PLOT = TRUE, # display plot
                      GAP = 0.1) # length of y-axis above and below max and min                                       

# Extract results
tau_values <- result$Value
p_values <- result$Test

# View tau coefficients
print(tau_values)
print(p_values)


# Prepare data frame with 3 columns: Date, Variable1, Variable2
# For auto-correlation, use same variable twice
marshalltown_con <- left_join(df_pds_marshalltown, marengo_daily, by = "Date") %>%
  select(Date, Flow.x, Flow.y) %>%
    rename(Marshalltown = Flow.x,
           Marengo = Flow.y) %>%
  na.omit()

head(marshalltown_con)

# Calculate Kendall's tau at different lags
result <- MultiHazard::Kendall_Lag(Data = marshalltown_con,
                      Lags = seq(-lag, lag, 1),  # lag is applied to second named variable
                      PLOT = TRUE, # display plot
                      GAP = 0.1) # length of y-axis above and below max and min                                     

# Extract results
tau_values <- result$Value
p_values <- result$Test

# View tau coefficients
print(tau_values)
print(p_values)
```

The maximum tau in both scenarios occurs with a lag of 0, meaning that the coincident events should be pulled from the same date as the extremal (PDS) events.

# Scenario A (Extremal Marshalltown Flows) Marginal Distributions

Scenario A is defined by extremal events (PDS) at the Marshalltown gage and the coincident events at the Marengo gage. The conditioning variable is Marshalltown flows, which are conditioned on Marengo flows.

Kendall's tau is max at a lag of 0 days. Therefore, use left_join to match coincident events to the date of the PDS events.

```{r}

ggplot(marshalltown_con, aes(x = Marshalltown, y = Marengo)) +
    geom_point() +
    xlab("Marshalltown PDS Flow (cfs)") + 
    ylab("Marengo Coincident Flow (cfs)") +
    ggtitle("Iowa River") +
    theme_bw()

# Fit marginal distributions
marshalltown_gpa <- lmom::pelgpa(lmom::samlmu(marshalltown_con$Marshalltown)) 
marshalltown_gev <- madsen_annualization(marshalltown_gpa, rate_marshalltown)
marengo_glo <- lmom::pelglo(lmom::samlmu(marshalltown_con$Marengo))

evplot(marshalltown_con$Marshalltown, main = "Marshalltown"); 
# evdistq(quagev, para = marshalltown_gev)

marshalltown_con_gev <- marshalltown_con %>%
  mutate(marshalltown_cdf = lmom::cdfgev(Marshalltown, marshalltown_gev),
         marengo_cdf = lmom::cdfglo(Marengo, marengo_glo))

marshalltown_con_gpa <- marshalltown_con %>%
  mutate(marshalltown_cdf = lmom::cdfgpa(Marshalltown, marshalltown_gpa),
         marengo_cdf = lmom::cdfglo(Marengo, marengo_glo))
```

# Scenario B (Extremal Marengo Flows) Marginal Distributions

Scenario B is defined by extremal events (PDS) at the Marengo gage and the coincident events at the Marshalltown gage. The conditioning variable is Marengo flows, which are conditioned on Marshalltown flows.

```{r}

ggplot(marengo_con, aes(x = Marengo, y = Marshalltown)) +
    geom_point() +
    xlab("Marengo PDS Flow (cfs)") + 
    ylab("Marshalltown Coincident Flow (cfs)") +
    ggtitle("Iowa River") +
    theme_bw()

# Fit marginal distributions
marengo_gpa <- lmom::pelgpa(lmom::samlmu(marengo_con$Marengo)) 
marengo_gev <- madsen_annualization(marengo_gpa, rate_marengo)
marshalltown_glo <- lmom::pelglo(lmom::samlmu(marengo_con$Marshalltown))

evplot(marshalltown_con$Marshalltown, main = "Marshalltown"); 
# evdistq(quagev, para = marshalltown_gev)

marengo_con_gev <- marengo_con %>%
  mutate(marengo_cdf = lmom::cdfgev(Marengo, marengo_gev),
         marshalltown_cdf = lmom::cdfglo(Marshalltown, marshalltown_glo))

marengo_con_gpa <- marengo_con %>%
  mutate(marengo_cdf = lmom::cdfgpa(Marengo, marengo_gpa),
         marshalltown_cdf = lmom::cdfglo(Marshalltown, marshalltown_glo))
```

# Scenario A Copula Fits and Comparisons

```{r}

u <- pull(marshalltown_con_gpa, marshalltown_cdf)
v <- pull(marshalltown_con_gpa, marengo_cdf)

gumbel_cop <- BiCopEst(u, v, family = 4)
print(gumbel_cop$AIC)
plot(gumbel_cop, type = "contour", margins = "unif", main = "Gumbel", col = "blue")
points(x=u, y=v, pch = 16)

t_cop <- BiCopEst(u, v, family = 2)
print(t_cop$AIC)
plot(t_cop, type = "contour", margins = "unif", main = "Student's t", col = "blue")
points(x=u, y=v, pch = 16)

gauss_cop <- BiCopEst(u, v, family = 1)
print(gauss_cop$AIC)
plot(gauss_cop, type = "contour", margins = "unif", main = "Gaussian", col = "blue")
points(x=u, y=v, pch = 16)

clayton_cop <- BiCopEst(u, v, family = 3)
print(clayton_cop$AIC)
plot(clayton_cop, type = "contour", margins = "unif", main = "Clayton", col = "blue")
points(x=u, y=v, pch = 16)

frank_cop <- BiCopEst(u, v, family = 5)
print(frank_cop$AIC)
plot(frank_cop, type = "contour", margins = "unif", main = "Frank", col = "blue")
points(x=u, y=v, pch = 16)

joe_cop <- BiCopEst(u, v, family = 6)
print(joe_cop$AIC)
plot(joe_cop, type = "contour", margins = "unif", main = "Joe", col = "blue")
points(x=u, y=v, pch = 16)

# Create summary table
copula_tbl <- dplyr::bind_rows(
  tibble(Copula = "Gumbel",       family = gumbel_cop$family,  par = gumbel_cop$par,  par2 = gumbel_cop$par2,  AIC = gumbel_cop$AIC),
  tibble(Copula = "Student's t",  family = t_cop$family,       par = t_cop$par,       par2 = t_cop$par2,       AIC = t_cop$AIC),
  tibble(Copula = "Gaussian",     family = gauss_cop$family,   par = gauss_cop$par,   par2 = gauss_cop$par2,   AIC = gauss_cop$AIC),
  tibble(Copula = "Clayton",      family = clayton_cop$family, par = clayton_cop$par, par2 = clayton_cop$par2, AIC = clayton_cop$AIC),
  tibble(Copula = "Frank",        family = frank_cop$family,   par = frank_cop$par,   par2 = frank_cop$par2,   AIC = frank_cop$AIC),
  tibble(Copula = "Joe",          family = joe_cop$family,     par = joe_cop$par,     par2 = joe_cop$par2,     AIC = joe_cop$AIC)
) %>%
  mutate(
    dep_parameter = ifelse(is.na(par2), 
                           sprintf("%.4f", par),
                           sprintf("par=%.4f, par2=%.4f", par, par2))
  ) %>%
  select(Copula, dep_parameter, AIC) %>%
  arrange(AIC)

copula_tbl

# Select the Copula with the minimum AIC (in this scenario, the Joe copula)
marshalltown_cop <- joe_cop
```

# Scenario B Copula Fits and Comparisons

```{r}

u <- pull(marengo_con_gpa, marengo_cdf)
v <- pull(marengo_con_gpa, marshalltown_cdf)

gumbel_cop <- BiCopEst(u, v, family = 4)
print(gumbel_cop$AIC)
plot(gumbel_cop, type = "contour", margins = "unif", main = "Gumbel", col = "blue")
points(x=u, y=v, pch = 16)

t_cop <- BiCopEst(u, v, family = 2)
print(t_cop$AIC)
plot(t_cop, type = "contour", margins = "unif", main = "Student's t", col = "blue")
points(x=u, y=v, pch = 16)

gauss_cop <- BiCopEst(u, v, family = 1)
print(gauss_cop$AIC)
plot(gauss_cop, type = "contour", margins = "unif", main = "Gaussian", col = "blue")
points(x=u, y=v, pch = 16)

clayton_cop <- BiCopEst(u, v, family = 3)
print(clayton_cop$AIC)
plot(clayton_cop, type = "contour", margins = "unif", main = "Clayton", col = "blue")
points(x=u, y=v, pch = 16)

frank_cop <- BiCopEst(u, v, family = 5)
print(frank_cop$AIC)
plot(frank_cop, type = "contour", margins = "unif", main = "Frank", col = "blue")
points(x=u, y=v, pch = 16)

joe_cop <- BiCopEst(u, v, family = 6)
print(joe_cop$AIC)
plot(joe_cop, type = "contour", margins = "unif", main = "Joe", col = "blue")
points(x=u, y=v, pch = 16)

# Create summary table
copula_tbl <- dplyr::bind_rows(
  tibble(Copula = "Gumbel",       family = gumbel_cop$family,  par = gumbel_cop$par,  par2 = gumbel_cop$par2,  AIC = gumbel_cop$AIC),
  tibble(Copula = "Student's t",  family = t_cop$family,       par = t_cop$par,       par2 = t_cop$par2,       AIC = t_cop$AIC),
  tibble(Copula = "Gaussian",     family = gauss_cop$family,   par = gauss_cop$par,   par2 = gauss_cop$par2,   AIC = gauss_cop$AIC),
  tibble(Copula = "Clayton",      family = clayton_cop$family, par = clayton_cop$par, par2 = clayton_cop$par2, AIC = clayton_cop$AIC),
  tibble(Copula = "Frank",        family = frank_cop$family,   par = frank_cop$par,   par2 = frank_cop$par2,   AIC = frank_cop$AIC),
  tibble(Copula = "Joe",          family = joe_cop$family,     par = joe_cop$par,     par2 = joe_cop$par2,     AIC = joe_cop$AIC)
) %>%
  mutate(
    dep_parameter = ifelse(is.na(par2), 
                           sprintf("%.4f", par),
                           sprintf("par=%.4f, par2=%.4f", par, par2))
  ) %>%
  select(Copula, dep_parameter, AIC) %>%
  arrange(AIC)

copula_tbl

# Select the Copula with the minimum AIC (in this scenario, the Joe copula)
marengo_cop <- joe_cop
```

# Copula Simulation and Transformation

## Generate Samples

Generate a large number of samples (N = 10\^7). Split the number of samples across the two conditional copula models in proportion to the number of events in each dataset.

```{r}
N <- 10^7
n_A <- nrow(marshalltown_con) # Number of coincident events in scenario A
n_B <- nrow(marengo_con) # Number of coincident events in scenario B
n1 <- round((N * n_A) / (n_A + n_B), 0)
n2 <- N - n1

# Simulate U(0,1)^2 from each fitted copula
s1 <- VineCopula::BiCopSim(n1, obj = marshalltown_cop)
s2 <- VineCopula::BiCopSim(n2, obj = marengo_cop)

# plot(head(s1, 1000))
```

## Transform to Original Scale using Quantile Functions

Transform the simulated uniforms to original units using quantile functions.

```{r}
# Scenario A
marshalltown_gev <- lmomco::vec2par(marshalltown_gev, type = "gev")
marengo_glo <- lmomco::vec2par(marengo_glo, type = "glo")

marshalltown_A <- quagev(s1[1,], marshalltown_gev)
marengo_A <- quaglo(s1[2,], marengo_glo)

copula_sample_A <- data.frame(Marshalltown = marshalltown_A, Marengo = marengo_A)

# Scenario B
marengo_gev <- vec2par(marengo_gev, type = "gev")
marshalltown_glo <- vec2par(marshalltown_glo, type = "glo")

marengo_B <- quagev(s2[1,], marengo_gev)
marshalltown_B <- quaglo(s2[2,], marshalltown_glo)

copula_sample_B <- data.frame(Marshalltown = marshalltown_B, Marengo = marengo_B)

# Combine samples from both scenarios; used in KDE 
copula_sample <- rbind(copula_sample_A, copula_sample_A)
```

# Isoline Derivation

## Generate Unit Square

```{r}
# Using "low" resolution here to reduce run times
x <- c(10^(-5), seq(999.9*10^(-5), 1-(1*10^(-6)), 10^(-4)))
y <- c(10^(-5), seq(999.9*10^(-5), 1-(1*10^(-6)), 10^(-4)))
grid <- list(x = x, y = y)
u <- expand.grid(x, y)
```

## Evaluate Copula CDF

Evaluate the copula CDF at each point on the grid.

```{r}
# Evaluates CDF of the specified parametric copula
uCopA <- VineCopula::BiCopCDF(u[,1], u[,2], marshalltown_cop) 
uCopB <- VineCopula::BiCopCDF(u[,1], u[,2], marengo_cop) 
```

## Determine Return Period

Determine the return period of each point on the unit square.

This code block defines the sampling frequency for daily data and a function to compute the return period surface matrix for an x, y grid in column blocks.

```{r}
mu <- 365.25 # Sampling frequency for daily data

# Computes return period surface matrix z on a grid of x and y values in blocks of columns
make_z_surface_blocked <- function(x, y, Cmat, mean_waiting_time, block = 500L){
  nx <- length(x); ny <- length(y)
  z <- matrix(NA_real_, nrow = nx, ncol = ny) # Pre-allocate output matrix z

  # Iterate over the y-axis in chunks
  for (j0 in seq.int(1L, ny, by = block)) {
    j1 <- min(j0 + block - 1L, ny)
    yb <- y[j0:j1]

    # denom is nx × nb (nb = block size)
    denom <- 1 - x - rep(yb, each = nx) + as.vector(Cmat[, j0:j1, drop=FALSE])
    z[, j0:j1] <- mean_waiting_time / matrix(denom, nrow = nx, ncol = length(yb))
  }
  z
}
```

### Scenario A Return Period

```{r}

# Mean waiting time is the average time, in days, between events
mean_waiting_time_A <- mu / rate_marshalltown 

# Reshape into a matrix: rows correspond to x, columns correspond to y
# IMPORTANT: expand.grid(x,y) varies x fastest, then y
Cmat <- matrix(uCopA, nrow = length(x), ncol = length(y), byrow = FALSE)

z  <- make_z_surface_blocked(x, y, Cmat, mean_waiting_time_A, block=200L)

rp_50 <- 50 # return period in years
xy50 <- grDevices::contourLines(x, y, z, levels = mu * rp_50)

if (length(xy50) == 0) stop("No contour found for RP = ", rp_50)

df_cl_A_50 <- do.call(rbind, lapply(seq_along(xy50), function(i){
  data.frame(id=i, level=xy50[[i]]$level, marshalltown_cdf=xy50[[i]]$x, marengo_cdf=xy50[[i]]$y, rp = rp_50)
}))


rp_100 <- 100
xy100 <- grDevices::contourLines(x, y, z, levels = mu * rp_100)

df_cl_A_100 <- do.call(rbind, lapply(seq_along(xy100), function(i){
  data.frame(id=i, level=xy100[[i]]$level, marshalltown_cdf=xy100[[i]]$x, marengo_cdf=xy100[[i]]$y, rp = rp_100)
}))
```

### Scenario B Return Period

```{r}

# Mean waiting time is the average time, in days, between events
mean_waiting_time_B <- mu / rate_marengo

# Reshape into a matrix: rows correspond to x, columns correspond to y
# IMPORTANT: expand.grid(x,y) varies x fastest, then y
Cmat <- matrix(uCopB, nrow = length(x), ncol = length(y), byrow = FALSE)

z  <- make_z_surface_blocked(x, y, Cmat, mean_waiting_time_B, block=200L)

rp_50 <- 50 # return period in years
xy50 <- grDevices::contourLines(x, y, z, levels = mu * rp_50)

if (length(xy50) == 0) stop("No contour found for RP = ", rp_50)

df_cl_B_50 <- do.call(rbind, lapply(seq_along(xy50), function(i){
  data.frame(id=i, level=xy50[[i]]$level, marshalltown_cdf=xy50[[i]]$x, marengo_cdf=xy50[[i]]$y, rp = rp_50)
}))


rp_100 <- 100
xy100 <- grDevices::contourLines(x, y, z, levels = mu * rp_100)

df_cl_B_100 <- do.call(rbind, lapply(seq_along(xy100), function(i){
  data.frame(id=i, level=xy100[[i]]$level, marshalltown_cdf=xy100[[i]]$x, marengo_cdf=xy100[[i]]$y, rp = rp_100)
}))
```

## Transform Variates to Original Scale

### Scenario A Transform Variates

```{r}

# Scenario A: Marshalltown PDS, Marengo coincident events

# Langbein adjustment defines the relationship between empirical flood expectancies in the partial duration series and the probability of the correspoding flood as an annual series
# p_pds is the partial duration non-exceedance probability, lambda is the rate of floods
# returns the annual non-exceedance probability
# Langbein, W. B. (1949). Annual Floods and the Partial-duration Flood Series. American Geophysical Union, 30(6), 879-881.
langbein_adjustment <- function(lambda, p_pds) {
  exp(-lambda * (1 - p_pds))
}

df_cl_A <- rbind(df_cl_A_50, df_cl_A_100)
p_pds <- df_cl_A$marshalltown_cdf
p_marshalltown_A <- langbein_adjustment(rate_marshalltown, p_pds)

df_cl_A <- df_cl_A %>%
  mutate(marshalltown_annualized_p = p_marshalltown_A,
         marshalltown = quagev(marshalltown_annualized_p, marshalltown_gev),
         marengo = quaglo(marengo_cdf, marengo_glo),
         scenario = "A")
```

### Scenario B Transform Variates

```{r}
df_cl_AB<- rbind(df_cl_B_50, df_cl_B_100)
p_pds <- df_cl_B$marengo_cdf
p_marengo_B <- langbein_adjustment(rate_marengo, p_pds)

df_cl_B <- df_cl_B %>%
  mutate(marengo_annualized_p = p_marengo_B,
         marengo = quagev(marengo_annualized_p, marengo_gev),
         marshalltown = quaglo(marshalltown_cdf, marshalltown_glo),
         scenario = "B")
```

## Plot Isolines

```{r}

ggplot(df_cl_A, aes(marshalltown, marengo, group = rp)) +
  geom_path(linewidth = 1) +
  theme_bw() +
  xlab("Marshalltown Flow (cfs)") +
  ylab("Marengo Flow (cfs)") +
  ggtitle("Scenario A: Extremal Marshalltown Flows") +
  xlim(0, 40000) +
  ylim(0, 40000) +
  coord_fixed(ratio = 1)

ggplot(df_cl_B, aes(marengo, marshalltown, group = rp)) +
  geom_path(linewidth = 1) +
  theme_bw() +
  xlab("Marengo Flow (cfs)") +
  ylab("Marshalltown Flow (cfs)") +
  ggtitle("Scenario B: Extremal Marengo Flows") +
  xlim(0, 50000) +
  ylim(0, 50000) +
  coord_fixed(ratio = 1)

df_cl_A_join <- df_cl_A %>%
  select(rp, marshalltown, marengo, scenario)

df_cl_B_join <- df_cl_B %>%
  select(rp, marshalltown, marengo, scenario)

df_cl <- rbind(df_cl_A_join, df_cl_B_join)

ggplot(df_cl, aes(marshalltown, marengo, color = scenario, linetype = factor(rp))) +
  geom_path(linewidth = 1) +
    scale_color_manual(values = c(
    "A" = "darkgreen",
    "B" = "purple")) +
  scale_linetype_manual(values = c(
    "50" = "dashed",
    "100" = "solid")) +
  xlim(0, 50000) +
  ylim(0, 50000) +
  theme_bw() +
  labs(color = "Scenario", linetype = "Return Period (years)") +
  xlab("Marshalltown Flow (cfs)") +
  ylab("Marengo Flow (cfs)") +
  ggtitle("") +
  coord_fixed(ratio = 1)

# View 1/100 joint probability contour from scenario A
df_cl[df_cl$scenario == "A" & df_cl$rp == "100", ]
```

# Combine Isolines

```{r}
combine_isolines_envelope <- function(
  iso_a, iso_b,
  x_col = "marshalltown",
  y_col = "marengo",
  step = NULL,          # e.g. 1, 0.1; if NULL uses unique rounded x/y
  digits = 2
){
  stopifnot(all(c(x_col, y_col) %in% names(iso_a)),
            all(c(x_col, y_col) %in% names(iso_b)))

  a <- iso_a[, c(x_col, y_col)]
  b <- iso_b[, c(x_col, y_col)]
  colnames(a) <- c("x","y")
  colnames(b) <- c("x","y")

  # build x grid
  if (is.null(step)) {
    x_grid <- sort(unique(round(c(a$x, b$x), digits)))
  } else {
    x_grid <- round(seq(min(c(a$x,b$x), na.rm=TRUE),
                        max(c(a$x,b$x), na.rm=TRUE),
                        by=step), digits)
  }

  # helper: max y at each x via interpolation
  max_y_at_x <- function(df, xg){
    df <- df[order(df$x),]
    yhat <- approx(df$x, df$y, xout=xg, rule=2)$y
    yhat
  }

  y_a <- max_y_at_x(a, x_grid)
  y_b <- max_y_at_x(b, x_grid)
  y_env <- pmax(y_a, y_b)
  part1 <- data.frame(x = x_grid, y = y_env)

  # build y grid
  if (is.null(step)) {
    y_grid <- sort(unique(round(c(a$y, b$y), digits)))
  } else {
    y_grid <- round(seq(min(c(a$y,b$y), na.rm=TRUE),
                        max(c(a$y,b$y), na.rm=TRUE),
                        by=step), digits)
  }

  # helper: max x at each y via interpolation
  max_x_at_y <- function(df, yg){
    df <- df[order(df$y),]
    xhat <- approx(df$y, df$x, xout=yg, rule=2)$y
    xhat
  }

  x_a <- max_x_at_y(a, y_grid)
  x_b <- max_x_at_y(b, y_grid)
  x_env <- pmax(x_a, x_b)
  part2 <- data.frame(x = x_env, y = y_grid)

  # combine, round, deduplicate (prefer unique x)
  out <- rbind(part1, part2)
  out$x <- round(out$x, digits)
  out$y <- round(out$y, digits)

  out <- out[order(out$x), ]
  out <- out[!duplicated(out$x), ]  # mimic your original "unique x" behavior
  out
}

iso_combined <- combine_isolines_envelope(df_cl_A, df_cl_B,
                                         x_col="marshalltown", y_col="marengo",
                                         step=1, digits=2)

ggplot(iso_combined, aes(x = x, y = y)) +
  geom_path(linewidth = 1.1, colour = "steelblue") +
  xlim(0, 50000) +
  ylim(0, 50000) +
  theme_bw() +
  xlab("Marshalltown Flow (cfs)") +
  ylab("Marengo Flow (cfs)") +
  ggtitle("") +
  coord_equal()
  
```

```{r}
kde_along_isoline <- function(Iso, cop_sample, Data, Isoline_Probs, Sim_Max){
  # truncate extreme simulated values (your original)
  remove <- which(cop_sample[,1] > Sim_Max*max(Data[,1], na.rm=TRUE) |
                  cop_sample[,2] > Sim_Max*max(Data[,2], na.rm=TRUE))
  if (Isoline_Probs == "Sample" && length(remove) > 1){
    cop_sample <- cop_sample[-remove, , drop=FALSE]
  }

  if (Isoline_Probs == "Sample"){
    pred <- ks::kde(x=cop_sample, eval.points=Iso)$estimate
  } else {
    pred <- ks::kde(x=stats::na.omit(Data), eval.points=Iso)$estimate
  }

  contour <- (pred - min(pred)) / (max(pred) - min(pred))
  list(prediction=pred, contour=contour)
}

extract_design_events_2d <- function(Iso, prediction, N_Ensemble){
  idx <- which(prediction == max(prediction, na.rm=TRUE))[1]
  most <- data.frame(Iso[idx,1], Iso[idx,2])

  # your current "full dependence" is max-x and max-y on the isoline (not true FD, but keep behavior)
  full <- data.frame(max(Iso[,1]), max(Iso[,2]))

  ens <- NULL
  if (N_Ensemble > 0){
    ok <- which(prediction > 0)
    ens <- Iso[sample(ok, size=N_Ensemble, replace=TRUE, prob=prediction[ok]), , drop=FALSE]
  }

  list(MostLikelyEvent=most, FullDependence=full, Ensemble=ens)
}
```
